Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-06-21 2017-Xudong-SMT/submit-ICTAC2017.tex
head of s5

% Macros for the SOS rules and proof trees:
%\newcommand\openrule[2]{\redinfer{#1}{#2}}
\newcommand\openrule[2]{\inferrule*[myfraction=\reddottedrule,center]{#1}{#2}}
%\newcommand\openrule[2]{\inferrule*{#1}{#2}}
%\newcommand\ostate[1]{\triangleleft{\;#1\;}\triangleright}
\newcommand\ostate[1]{\triangleleft{#1}\triangleright}
\newcommand{\sm}[1]{\mbox{\boldmath\small #1}}
\usepackage{algorithm} 
%\usepackage{algorithmicx} 
\usepackage{algpseudocode}  
\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\algnewcommand{\IfThenElse}[3]{\State \algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}% \IfThenElse{<if>}{<then>}{<else>}

@@

A composite pNet consists of a set of sub-pNets exposing
a set of actions, each of them triggering internal actions in each of
the sub-pNets. The synchronization between global actions and
internal actions is given by  \emph{synchronization vectors}: a
synchronization vector synchronizes one or several internal actions, and
exposes a single resulting global action.
% Actions involved at the pNet level (in the synchronization vectors) do
% not need to distinguish between input and output
% variables.
Action terms for pNets are defined as follows:
\[\begin{array}[l]{rcl@{\quad}l}
  \alpha\in \AlgAS &::=&a(Expr_1,\ldots,Expr_n)
\end{array}
\]

@@

\begin{example}\emph{An open-transition.}
  \label{OT:lotos-composed}

The \texttt{LOTOS} pNet of Fig. \ref{schema:lotos-pnet} has 2
controllers and 2 holes. We will show its full open Automaton in
section \ref{section:full-result}. We give one of its transitions here:

 \smallskip
 $ ot_2  = \openrule{
                       \{0 \xrightarrow{\delta}_{C_1} 0, 0 \xrightarrow{acc(x)}_{C_2} 1\}, ~~
                       \{\xrightarrow{hb_{12}}_P\}, ~~ 
                        [s_0=0 \!\land\! hb_{12}=\delta(x)], ~~ 
                       \{s_0 \xleftarrow{} 1\}     
                      }
     {\ostate{00} \xrightarrow{\underline{\delta(x)}} \ostate{01}}$
\end{example}

The open transition $ot_2$ involves the possible actions from the process $P$ and prefix expression $acc(x);Q$. The hole behavior, represented by variable $hb_{12}$ must be equal to the exit action $\delta(x)$, and the accept event $acc(x)$ will capture the same value $x$.

The transition in the conclusion of this rule starts in
state $\ostate{00}$, where both $C_1$ and $C_2$ are in their initial
states. It ends in a different structured state, because in $ot_2$
both pLTS conduct transitions, showing that the control has gone from
processes $P$ to $(acc(x);Q$, consuming actions $\delta(x)$ and
$acc(x)$. Controller $C_1$ has only one state, but its state variable
$s_0$ has changed to 1.

@@

\QIN{
For the open pNets, in previous work, we already have a sketch for computing the open automaton.
}

\QIN{
1) Start with the initial state of the automaton. 
}

\QIN{
2) Apply Tr1 and Tr2 to deduct all the possible open transitions. Use SMT solver to prune the 
unsatisfiable transitions. Recursively conduct this step with the target state of the transitions 
until all the global states are explored.
}

\QIN{
3) Finally a reachability check could potentially prune the state space.
}

However, more details are needed when doing the implementation. 
Algorithm 1 shows there are much more things added, such as how we apply 
the Tr1 or Tr2.

@@

\State Initialize a set for unexplored global states $U$, a set for explored global states $E$;
\State Initialize the initial global state $G$ of the open automaton, add it into $U$;
\State Initialize a list $L$ for storing open transitions;
\While{!isEmpty(U)}
	\State Extract a global state $S$ from the $U$;
	\State For the pNet node, recursively applying Tr1 or Tr2 on its subnets;
	\If{Subnet is pLTS, apply the Tr1} 
		\State Generate open transition $OT$ according to the transitions in pLTS;
	\EndIf
	\If{Subnet is pNet, apply the Tr2} 
		\For{each $sv \in SV$}
			\State Invoke $Combining()$ to generate the combination of sub-transitions;
			\State Invoke $Matching()$ to match the combination with the $sv$;
			\State Generate open transition $OT$ using the predicate from the former steps;
		\EndFor
	\EndIf
	\State Store $OT$ into $L$;
	\State Get the target global state $T$ from $OT$;
	\If{$(U.contains(T))\ ||\ (E.contains(T))$} 
		\State Add $T$ into the $U$;
	\EndIf
\EndWhile
\State Prune the open transitions in $L$ using SMT solver;
\State \Return $L$;

@@

>>>>

% Macros for the SOS rules and proof trees:
%\newcommand\openrule[2]{\redinfer{#1}{#2}}
\newcommand\openrule[2]{\inferrule*[myfraction=\reddottedrule,center]{#1}{#2}}
%\newcommand\openrule[2]{\inferrule*{#1}{#2}}
%\newcommand\ostate[1]{\triangleleft{\;#1\;}\triangleright}
\newcommand\ostate[1]{\triangleleft{#1}\triangleright}
\newcommand{\sm}[1]{\mbox{\boldmath\small #1}}
\usepackage{algorithm} 
%\usepackage{algorithmicx} 
\usepackage{algpseudocode}  
\renewcommand{\algorithmicrequire}{\textbf{Input:}} 
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\algnewcommand{\IfThen}[2]{\State \algorithmicif\ #1\ \algorithmicthen\ #2}% \IfThenElse{<if>}{<then>}
\algnewcommand{\IfThenElse}[3]{\State \algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}% \IfThenElse{<if>}{<then>}{<else>}

@@

A composite pNet consists of a set of sub-pNets exposing
a set of actions, each of them triggering internal actions in each of
the sub-pNets. The synchronization between global actions and
internal actions is given by  \emph{synchronization vectors}: a
synchronization vector synchronizes one or several internal actions, and
exposes a single resulting global action.
% Actions involved at the pNet level (in the synchronization vectors) do
% not need to distinguish between input and output
% variables.
%Action terms for pNets are defined as follows:
%\[\begin{array}[l]{rcl@{\quad}l}
%  \alpha\in \AlgAS &::=&a(Expr_1,\ldots,Expr_n)
%\end{array}
%\]

@@

%\begin{example}\emph{An open-transition.}
%  \label{OT:lotos-composed}
%
%The \texttt{LOTOS} pNet of Fig. \ref{schema:lotos-pnet} has 2
%controllers and 2 holes. We will show its full open Automaton in
%section \ref{section:full-result}. We give one of its transitions here:
%
% \smallskip
% $ ot_2  = \openrule{
%                       \{0 \xrightarrow{\delta}_{C_1} 0, 0 \xrightarrow{acc(x)}_{C_2} 1\}, ~~
%                       \{\xrightarrow{hb_{12}}_P\}, ~~ 
%                        [s_0=0 \!\land\! hb_{12}=\delta(x)], ~~ 
%                       \{s_0 \xleftarrow{} 1\}     
%                      }
%     {\ostate{00} \xrightarrow{\underline{\delta(x)}} \ostate{01}}$
%\end{example}
%
%The open transition $ot_2$ involves the possible actions from the process $P$ and prefix expression $acc(x);Q$. The hole behavior, represented by variable $hb_{12}$ must be equal to the exit action $\delta(x)$, and the accept event $acc(x)$ will capture the same value $x$.
%
%The transition in the conclusion of this rule starts in
%state $\ostate{00}$, where both $C_1$ and $C_2$ are in their initial
%states. It ends in a different structured state, because in $ot_2$
%both pLTS conduct transitions, showing that the control has gone from
%processes $P$ to $(acc(x);Q$, consuming actions $\delta(x)$ and
%$acc(x)$. Controller $C_1$ has only one state, but its state variable
%$s_0$ has changed to 1.

@@

@@

\State Initialize a set for unexplored global states $U$, a set for explored global states $E$;
\While{!isEmpty(U)}
	\State Extract a global state $S$ from the $U$;
	\State Start at the root pNet node
	\State Recursively applying Tr1 or Tr2 on it and its subnets;
	\If{Subnet is pLTS, apply the Tr1} 
		\For{each transition $TR$ start from the state $s$ in $S$}
			\State Generate open transition $OT$ according to $TR$;
			\State Store $OT$ into the result list $L$;
		\EndFor
	\EndIf
	\If{Subnet is pNet, apply the Tr2} 
		\For{each $sv \in SV$}
			\State Invoke $Combining()$ to generate the combination of sub-transitions;
			\State Invoke $Matching()$ to match the combination with the $sv$;
			\State Store the result set of open transitions into $L$;
		\EndFor
	\EndIf
	\For{each target global state $T$ from $OT \in L$}
		\IfThen{$(!U.contains(T))\ \&\&\ (!E.contains(T))$}{Add $T$ into the $U$;}
	\EndFor
\EndWhile
\State Prune the open transitions in $L$ using SMT solver;
\State \Return $L$;

@@

\QIN{
For the open pNets, in previous work, we already have a sketch for computing the open automaton.
The computation repeats applying Tr1 and Tr2 to deduct all the possible open transitions and pruning the 
unsatisfiable transitions by SMT solver until all the global states are explored. 
In the end, a reachability check is supposed to potentially prune the state space. 
Algorithm 1 is what we used in the implementation, there are much more details added based on 
the sketch, such as how we apply the Tr1 or Tr2.
}
